---
title: Stock Optimization
output: github_document
---
## Stock Portfolio Optimization
We will look at the S&P/TSX Index top performance stock names and compare it to the performance of Technology Index XLK.
The list for stocks is taken from an HTML table from Wikipedia and CAPM analysis is done using Yahoo Finance API.

# 1. Initial Set Up and Demonstration
```{r}
library(tidyverse)
library(tidyquant)
# Set random seed for reproducibility reasons

set.seed(301)

#Get HTML tables containing all stocks in TSX composite index,
#number 2 to refer the table index

ticker_names_table <-
    htmltab::htmltab("https://en.wikipedia.org/wiki/List_of_S%26P_600_companies",1)

#Get all stock tickers from column Ticker using pipe operator 
#%>% passing from left to right
ticker_names <- as.vector(ticker_names_table %>% select("Symbol"))
show(ticker_names)
```

Using Yahoo finance requires specific format, especially for stocks with period. We need to transform it into a hyphen
#and append ".TO" at the end to each tickers. Update: New version of tidyquant no longer requires this
```{r}
# ticker_names_transform <- gsub("[[:punct:]]","-",ticker_names$Ticker)
# #Use stringi to parse and join string to our tickers
# ticker_names_transform <- as.vector(stringi::stri_join(ticker_names_transform, ".TO"))

#Now, using log returns to see the change in price with respect to the day before using tq_get from
#tidyquant
TSX_returns <- ticker_names$Symbol %>%
    tq_get(get = "stock.prices",
           from = "2023-01-01",
           to = "2023-07-01") %>%
    group_by(symbol) %>% #See periodReturn per ticker
    tq_transmute(select = adjusted,
                 mutate_fun = periodReturn, #mutate_fun create new columns using specific function, here finding the period return
                 period="daily",
                 type="log",
                 col_rename = "Ra") #Ra stands for return group a
show(TSX_returns)
```
Next, we will benchmark the stock against particular industry index in S&P500. We'll choose XLK here since
tq_get only works in the US Index

```{r}
XLK_returns <- "XLK" %>%
    tq_get(get= "stock.prices",
           from = "2023-01-01",
           to = "2023-07-01") %>%
    tq_transmute(select = adjusted,
                 mutate_fun =  periodReturn,
                 type = "log",
                 period = "daily",
                 col_rename = "Rb") #Return for group b 

#Perform left-join on Ra and Rb to see how well stocks perform against index
RaRb <- left_join(TSX_returns,
                  XLK_returns,
                  by = "date") %>%
        group_by(symbol)
show(RaRb)
```


```{r}
#Capital Asset Pricing Model (CAPM) : Expected Return R = alpha*Return - beta*Risk
CAPM <- RaRb %>%
        tq_performance(
            Ra=Ra,
            Rb = Rb,
            performance_fun = table.CAPM)

#Depend on return(Alpha) or Risk(Beta) preference, pick appropriate 
#indicator or Information Ratio a/b for both, or top quintile performance.
top_percentile <- CAPM %>%
                filter(AnnualizedAlpha > quantile(CAPM$AnnualizedAlpha, 0.9))

#Using the new tickers, return to our original log-return table to get its log-return
quintile_ticker = top_percentile$symbol

stock_returns_quintile <- TSX_returns %>% filter(symbol %in% quintile_ticker)
show(stock_returns_quintile)
```
## Performing Portfolio Distribution
# 1. Single case method
We need to manually set weights since tq_portfolio doesn't have filter function
```{r}
set.seed(5)
weights <- c(1:100)
random_weights <- sample(weights, size=length(quintile_ticker), replace = TRUE)
index_for_zeroes <- sample(c(1:length(random_weights)),round(length(quintile_ticker)*0.78,0)) #37 out of 47 Tickers will be 0
random_weights[index_for_zeroes] <- 0 #Insert zeroes accordingly
#Random weigth of 0 yields result similar to Gaussian Hyper Parameter Search, which is better than quadratic optimization in term of speed and performance

#Now, we need to ensure all weights sum up as 1 by normalizing them
weights_mapping_single <- tibble(
    symbols = quintile_ticker,
    weights = random_weights) %>%
        mutate( weights = weights/sum(weights))

#Use tq_portfolio to pick portfolio with non-zero weights. Respective return = sum of product of log return X weights
portfolio_returns_daily <- stock_returns_quintile %>% 
    tq_portfolio(assets_col = symbol,
                 returns_col =  Ra,
                 weights =  weights_mapping_single,
                 col_rename = "Ra")

#Now, we join Benchmark return to our portfolio
RaRb_single_portfolio <- left_join(portfolio_returns_daily,
                                   XLK_returns,
                                   by = "date")
show(RaRb_single_portfolio)
```
Then, we continue with CAPM analysis 
```{r}
#Perform CAPM analysis
RaRb_single_portfolio %>%
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = table.CAPM)
RaRb_single_portfolio %>%
  tq_performance(Ra = Ra, Rb = Rb, performance_fun = InformationRatio)
```
Finally, a bar plot should help us visualize the performance
```{r}

```