---
title: Stock Optimization
output: github_document
---
## Stock Portfolio Optimization
We will look at the S&P/TSX Index top performance stock names and compare it to the performance of Technology Index XLK.
The list for stocks is taken from an HTML table from Wikipedia and CAPM analysis is done using Yahoo Finance API.

# 1. Initial Set Up and Demonstration
```{r}
library(tidyverse)
library(tidyquant)
# Set random seed for reproducibility reasons

set.seed(301)

#Get HTML tables containing all stocks in TSX composite index,
#number 2 to refer the table index

ticker_names_table <-
    htmltab::htmltab("https://finance.yahoo.com/quote/QQQ/",-1)
ticker_names_table
#Get all stock tickers from column Ticker using pipe operator 
#%>% passing from left to right
#ticker_names <- as.vector(ticker_names_table %>% select("Symbol"))
#show(ticker_names)
```

Using Yahoo finance requires specific format, especially for stocks with period. We need to transform it into a hyphen
#and append ".TO" at the end to each tickers. Update: New version of tidyquant no longer requires this
# ```{r}
# # ticker_names_transform <- gsub("[[:punct:]]","-",ticker_names$Ticker)
# # #Use stringi to parse and join string to our tickers
# # ticker_names_transform <- as.vector(stringi::stri_join(ticker_names_transform, ".TO"))

# #Now, using log returns to see the change in price with respect to the day before using tq_get from
# #tidyquant
# TSX_returns <- ticker_names$Symbol %>%
#     tq_get(get = "stock.prices",
#            from = "2023-01-01",
#            to = "2023-07-01") %>%
#     group_by(symbol) %>% #See periodReturn per ticker
#     tq_transmute(select = adjusted,
#                  mutate_fun = periodReturn, #mutate_fun create new columns using specific function, here finding the period return
#                  period="daily",
#                  type="log",
#                  col_rename = "Ra") #Ra stands for return group a
# show(TSX_returns)
```

```{r}
#Update date everytime we run
current_date = Sys.Date()
YOY_date = current_date-years(1)

TSX_returns <- 
    tq_get("QQQ",get = "stock.prices",
           from = YOY_date,
           to = current_date) %>%
    group_by("QQQ") %>% #See periodReturn per ticker
    tq_transmute(select = adjusted,
                 mutate_fun = periodReturn, #mutate_fun create new columns using specific function, here finding the period return
                 period="daily",
                 type="log",
                 col_rename = "Ra_QQQ") %>% arrange(desc(date)) #Ra stands for return group a 
show(TSX_returns)

```

Next, we will benchmark the stock against particular industry index in S&P500. We'll choose S&P500 here since
tq_get only works in the US Index

```{r}
XLK_returns <- "^GSPC" %>%
    tq_get(get= "stock.prices",
           from = YOY_date,
           to = current_date) %>%
    #group_by("XLK") %>%
    tq_transmute(select = adjusted,
                 mutate_fun =  periodReturn,
                 type = "log",
                 period = "daily",
                 col_rename = "Rb_XLK") #Return for group b 

#Perform left-join on Ra and Rb to see how well stocks perform against index
RaRb <- left_join(TSX_returns,
                  XLK_returns,
                  by = "date")

show(RaRb)

```

#Perform CAPM analysis

```{r}
#Use tq_portfolio to pick portfolio with non-zero weights. Respective return = sum of product of log return X weights
portfolio_returns_daily <- RaRb 

portfolio_returns_daily %>%
  tq_performance(Ra = Ra_QQQ, Rb = Rb_XLK, performance_fun = table.CAPM)
portfolio_returns_daily %>%
  tq_performance(Ra = Ra_QQQ, Rb = Rb_XLK, performance_fun = InformationRatio)

#Plot daily returns using bar plot with regresion line for trend

portfolio_returns_daily %>%
  ggplot(aes(x = date, y = Ra_QQQ)) +
  geom_bar(stat = "identity", fill = palette_light()[[1]]) +
  labs(title = "Portfolio Returns",
       x = "", y = "Monthly Returns") +
  geom_smooth(method = "lm") +
  theme_tq() +
  scale_color_tq() +
  scale_y_continuous(labels = scales::percent)
```
```{r}
#Test out our portfolio with 10,000 investment.
#Use tq_portfolio to pick portfolio with non-zero weights. Respective return = sum of product of log return X weights
portfolio_returns_daily <- TSX_returns %>%
    tq_portfolio(assets_col = "QQQ",
                 returns_col =  Ra_QQQ,
                 col_rename = "Ra_QQQ"
                 )

symbol_a = c(rep('PRT', length(portfolio_returns_daily$Ra_QQQ)))

portfolio_asset_name = cbind(portfolio_returns_daily, symbol_a)

portfolio_growth_single <- portfolio_asset_name %>%
    tq_portfolio(assets_col   = symbol_a, 
                returns_col  = Ra_QQQ, 
                col_rename   = "investment.growth",
                wealth.index = TRUE) %>%
    mutate(investment.growth = investment.growth * 10000)

#Now, let's compare it with our bask XLK to see how our portfolio return
symbol_b = c(rep('XLK', length(XLK_returns$Rb_XLK)))

XLK_asset_name = cbind(XLK_returns, symbol_b)

portfolio_growth_XLK <- XLK_asset_name %>%
    tq_portfolio(assets_col   = symbol_b, 
                returns_col  = Rb_XLK, 
                col_rename   = "investment.growth",
                wealth.index = TRUE) %>%
    mutate(investment.growth = investment.growth * 10000)

#Plot both portfolio, with blue being our pick and red being the index fund
ggplot() + 
geom_line(data = portfolio_growth_single, aes(x = date, y = investment.growth), color = "blue") +
geom_line(data = portfolio_growth_XLK, aes(x = date, y = investment.growth), color = "red") +
labs(title = "Portfolio Growth", x = "", y = "Portfolio Value") +
theme_tq() +
scale_color_tq() +
scale_y_continuous(labels = scales::dollar)
```

test <- bind_rows(list("QQQ"=portfolio_growth_single,"SP500"=portfolio_growth_XLK),.id="Portfolio")
#Plot both portfolio, with blue being our pick and red being the index fund
test %>% 
  ggplot() + 
  geom_line(data = test, aes(x = date, y = investment.growth, color = Portfolio)) +
  labs(title = "Portfolio Growth", x = "", y = "Portfolio Value") +
  theme_tq() +
  scale_color_tq() +
  scale_y_continuous(labels = scales::dollar)
